{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-10-08 thesis/future.tex
try not to aggro garth so much abt ml
change future quote

\chapter{Future Work}
\label{chap:warpzone}
% TODO: ask for permission for this one
% https://twitter.com/Poem4your_sprog/status/878747092788838400
\inspirationalquote{
\begin{tabular}{p{0.51\textwidth}}
A little bit of being kind,
a tiny open door; \\
A nicer slice of peace of mind
begets a little more. \\
It won't correct the world tonight,
nor change tomorrow too; \\
But maybe if you do it right,
you'll find it changes you.
\end{tabular}}
{Sam "Poem\_for\_your\_sprog" Garland}

@@

In my experience,
the WBE algorithm (used in Landslide for estimating number of interleavings) consistently underestimates,
being often seen to count (nearly) monotonically up toward the true value as exploration progresses,
while the RE algorithm (used in Landslide for estimating total execution time) can be unstable and erratic,
bouncing wildly among orders of magnitude even in the space of ten or fewer adjacent interleavings.
These flaws are especially visible in the {\tt avl\_fixed} transactional verification results (Table~\ref{tab:tm-verifs}).
Both estimation algorithms
use the known structure of the tree as a model for the unknown
(see \sect{\ref{sec:landslide-estimate}}),
but make no use of domain-specific knowledge such as which threads were chosen to run at each preemption point.
For example, consider the root of the left subtree in Figure~\ref{fig:tree},
thread 1's \dporTA{{\tt tmp1++}}.
At that moment thread 1 has 1 more instruction left to execute, while thread 2 has 3.
If thread 1's last instruction is chosen to run next,
there can only be one way to run thread 2's 3 steps thereafter (${3+0 \choose 0} = 1$);
if thread 2's first instruction is chosen to run next,
there will be 3 ways to interleave thread 1's last among thread 2's remaining 2 (${2+1 \choose 1} = 3$).
In general, the largest child subtree at any preemption point will be the one resulting from running the thread with the most transitions left,
or more formally, fixing $N+M=C$, ${N+M \choose N}$ is maximized when $N \approx M$.
WBE and RE could both be relatively easily extended to incorporate this observation,
using knowledge from branches already tested to guess (still heuristically) appropriate values for $N$ and $M$.
Predicting more advanced aspects of state space structure,
such as subtrees pruned by sleep set reduction
(\sect{\ref{sec:landslide-sleepsets}}; $\bigstar$4 in Figure~\ref{fig:dpor-example-2}),
would likely require analyzing DPOR's memory conflicts as well.

@@

% arugment why ML is NOT appropriate for smc in general -- if you try to train it to choose thread interleavings / search ordering,
% even if you're already in not-enough-cpu-time heuristics land, you're biasing it to find only the types of bugs you've already seen
% and destroying its ability to find novelties
% if your ML provides a partial verification, it is impossible for a human to understand what that verification means
As a matter of personal philosophy,
I would strongly caution against using machine learning to try to expose bugs faster.
Machine learning has become popular recently as a magic cure-all for many programming problems,
but is also notorious for amplifying any biases in its training set in often unpredictable ways
\cite{conceptnet-bias}.
In the domain of concurrency testing, this would translate to being able to find only types of bugs
already seen in the training data, abandoning any novel bug patterns to further obscurity.
Granted,
Iterative Deepening and
ICB also deprioritize ``deeper'' bugs,
as measured by number of unique or total preemptions, respectively,
%in terms of minimum necessary preemption count,
%but this strategy is informed by empirical evidence
%% TODO: upgrade citation to also have a section reference to that icb bound table from quicksand eval
%\cite{chess-icb},
%and
but when a bug is not found
these strategies are also transparent enough that the user can understand the nature of the partial verification.
%
Machine learning would be more appropriate for improving state space estimation,
whose output is already a black-box oracle from the user's perspective,
and where the worst case of being wrong
is that Quicksand prioritize its jobs suboptimally. % subjunctive
In-progress estimates fluctuate depending on many factors,
such as interleaving-dependent flow control,
and I suspect deeper patterns may exist among multiple different state spaces,
for example,
from the same test program with different $K,N$ parameters.
For any future researchers who wish to study such patterns,
% with machine learning... don't feel like i can do justice to discussing its flaws with real citations
I have made the estimation logs from the HTM experiments (\sect{\ref{sec:tm-eval}})
freely available at \url{https://github.com/bblum/thesis-htm-logs}.
%
Nevertheless,
any overall search heuristic for concurrency testing
must be designed to support the human user's agency first and foremost,
that she may readily understand partial verification results
and participate effectively in the test case refinement strategies discussed
upchapter. % this is a word now.

>>>>

\chapter{Future Work}
\label{chap:warpzone}
\inspirationalquote{
\begin{tabular}{p{0.75\textwidth}}
We can't be free until we learn to laugh at ourselves.
Once you look in the mirror and see just how foolish we can be,
laughter is inevitable.
And from laughter comes wisdom.
\end{tabular}}
{G'Kar, Babylon 5}

@@

% TODO: does this need some estimate progress graphs
% some RE ones would even need a log scale y axis maybe
In my experience,
the WBE algorithm (used in Landslide for estimating number of interleavings) consistently underestimates,
being often seen to count (nearly) monotonically up toward the true value as exploration progresses,
while the RE algorithm (used in Landslide for estimating total execution time) can be unstable and erratic,
bouncing wildly among orders of magnitude even in the space of ten or fewer adjacent interleavings.
These flaws are especially visible in the {\tt avl\_fixed} transactional verification results (Table~\ref{tab:tm-verifs}).
Both estimation algorithms
use the known structure of the tree as a model for the unknown
(see \sect{\ref{sec:landslide-estimate}}),
but make no use of domain-specific knowledge such as which threads were chosen to run at each preemption point.
For example, consider the root of the left subtree in Figure~\ref{fig:tree},
thread 1's \dporTA{{\tt tmp1++}}.
At that moment thread 1 has 1 more instruction left to execute, while thread 2 has 3.
If thread 1's last instruction is chosen to run next,
there can only be one way to run thread 2's 3 steps thereafter (${3+0 \choose 0} = 1$);
if thread 2's first instruction is chosen to run next,
there will be 3 ways to interleave thread 1's last among thread 2's remaining 2 (${2+1 \choose 1} = 3$).
In general, the largest child subtree at any preemption point will be the one resulting from running the thread with the most transitions left,
or more formally, fixing $N+M=C$, ${N+M \choose N}$ is maximized when $N \approx M$.
WBE and RE could both be relatively easily extended to incorporate this observation,
using knowledge from branches already tested to guess (still heuristically) appropriate values for $N$ and $M$.
Predicting more advanced aspects of state space structure,
such as subtrees pruned by sleep set reduction
(\sect{\ref{sec:landslide-sleepsets}}; $\bigstar$4 in Figure~\ref{fig:dpor-example-2}),
would likely require analyzing DPOR's memory conflicts as well.

@@

% arugment why ML is NOT appropriate for smc in general -- if you try to train it to choose thread interleavings / search ordering,
% even if you're already in not-enough-cpu-time heuristics land, you're biasing it to find only the types of bugs you've already seen
% and destroying its ability to find novelties
% if your ML provides a partial verification, it is impossible for a human to understand what that verification means

%% commenting this first part out bc it will probably aggro garth...
%As a matter of personal philosophy,
%I would strongly caution against using machine learning to try to expose bugs faster.
Machine learning has become popular recently as a magic cure-all for many programming problems,
but is also notorious for amplifying any biases in its training set in often unpredictable ways
\cite{conceptnet-bias}.
In the domain of concurrency testing, this would translate to being able to find only types of bugs
already seen in the training data, abandoning any novel bug patterns to further obscurity.
Granted,
Iterative Deepening and
ICB also deprioritize ``deeper'' bugs,
as measured by number of unique or total preemptions, respectively,
%in terms of minimum necessary preemption count,
%but this strategy is informed by empirical evidence
%% TODO: upgrade citation to also have a section reference to that icb bound table from quicksand eval
%\cite{chess-icb},
%and
but when a bug is not found
these strategies are also transparent enough that the user can understand the nature of the partial verification.
%
Machine learning would be more appropriate for improving state space estimation,
whose output is already a black-box oracle from the user's perspective,
and where the worst case of being wrong
is that Quicksand prioritize its jobs suboptimally. % subjunctive
In-progress estimates fluctuate depending on many factors,
such as interleaving-dependent flow control,
and I suspect deeper patterns may exist among multiple different state spaces,
for example,
from the same test program with different $K,N$ parameters.
For any future researchers who wish to study such patterns,
% with machine learning... don't feel like i can do justice to discussing its flaws with real citations
I have made the estimation logs from the HTM experiments (\sect{\ref{sec:tm-eval}})
freely available at \url{https://github.com/bblum/thesis-htm-logs}.
%
Nevertheless,
any overall search heuristic for concurrency testing
must be designed to support the human user's agency first and foremost,
that she may readily understand partial verification results
and participate effectively in the test case refinement strategies discussed
upchapter. % this is a word now.
